---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# fishLengthAssess

<!-- badges: start -->
<!-- badges: end -->

This repository contains the code and documentation for the fishLengthAssess R package. fishLengthAssess acts as an aggregator of length-based stock assessment models. The package imposes a standardized formatting and metadata structure on length data sets, which can then be used in a variety of length-based assessment methods that are aggregated within the package. It also provides wrapper functions to link standardized formatting of length data sets with external R packages that contain length-based assessment methods (e.g. LBSPR package).

## Installation

You can install the development version of fishLengthAssess from [GitHub](https://github.com/) with:
```{r, eval=FALSE}
# install.packages("devtools")
devtools::install_github("natureanalytics-ca/fishLengthAssess")
```

## User guide

Documentation for the fishLengthAssess package can be found [here](https://natureanalytics-ca.github.io/fishLengthAssess/) as a user guide.





