---
title: "fishLengthAssess: aggregator of length-based stock assessment methods"
author: "Ana Adao"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Example vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction


## Load the package and example data sets

Explain LifeHistoryObj and LengthCompObj

```{r setup, eval=FALSE}
library(fishLengthAssess)
```

# Main functions

### `PmatFunc` function

Description of this function...

### `PoptFunc` function

Description of this function...

### `PmegaFunc` function

Description of this function...

### `PLcFunc` function

Description of this function...

### `lbsprWrapper` function

Description of this function...


## Utility functions

### `poolLengthComp` function

Description of this function...

### `LoptFunc` function

Description of this function...

### `LcFunc` function

Description of this function...

